# startup
```{r}
library(tidyverse)
library(igraph)
library(visNetwork)
library(MASS)

# read file
nd <- as.data.frame(read_csv("./[INSERT_FILE_NAME].csv"))
# users sending
from_id <- nd[,1]
# users receiving
to_id <- nd[,2]
```

# add 1 to from_id list and to_id list
```{r}
# create new from_id list where we add 1 to each node id (this is because you cant have a node 0, must start at node 1)
# initialize from_id_plus list 
from_id_plus <- c()
# iterate through nodes in from_id list
for (i in from_id) {
  # add node id plus 1 to from_id_plus list
  from_id_plus <- c(from_id_plus, i+1)
}
# create new to_id list where we add 1 to each node id (this is because you cant have a node 0, must start at node 1)
# initialize to_id_plus list
to_id_plus <- c()
# iterate through nodes in to_id list
for (i in to_id) {
  # add node id plus 1 to to_id_plus list
  to_id_plus <- c(to_id_plus, i+1)
}
```

# create directed list connecting from and to nodes
```{r}
# list with nodes
from_to <- c()
# index
j <- 1
# for loop to generate nodes with directed edges
for (i in 1:dim(nd)) {
  # get users sending (need to add 1 as can't have nodes starting with 0 in R)
  from_to[j] <- nd[i,1]+1
  j <- j + 1
  # get users receiving
  from_to[j] <- nd[i,2]+1
  j <- j + 1
}
```



### GRAPHING

# create graph objects
```{r}
# create graph object
g <- graph(from_to)
# apply Louvain algorithm on graph object
new_g <- cluster_louvain(as.undirected(g, mode = "collapse"), weights = NULL, resolution = 1)
```

# create variables used to create node/edge data frame
```{r}
# initialize list to get number of elements in each group
group_lengths <- c()
# get all unique node ids
unique_node_ids <- c()
# iterate through nodes in new_g
for(i in 1:length(new_g)){
  # add group of nodes to unique_node_ids list
  unique_node_ids <- c(unique_node_ids, c(new_g[[i]]))
  # add group length to group_lengths list
  group_lengths <- c(group_lengths, length(new_g[[i]]))
}
```

# create node data frame variables (for visNetwork graph)
```{r}
# create node labels
# initialize node labels list
node_labels <- c()
# iterate through unique node ids
for (i in unique_node_ids) {
  # add node to node_labels
  node_labels <- c(node_labels, as.character(i))
}
# create node groups
# initialize node groups list
node_groups <- c()
# temp counter
j <- 1
# iterate through group_lengths list
for (i in group_lengths) {
  # repeat group id by value of i and add to node_groups list
  node_groups <- c(node_groups, rep(j, i))
  # increase j
  j <- j + 1
}
# create node titles
# initialize node titles list
node_titles <- c()
# iterate through unique node ids
for (i in unique_node_ids) {
  # create node title and add to list
  node_titles <- c(node_titles, paste("Node", as.character(i), "info"))
}
# create nodes data frame for Louvain algorithm applied graph
new_g_nodes <- data.frame(
  id = unique_node_ids,
  label = node_labels,
  group = node_groups,
  title = node_titles
)
```

# create edge data frame variables (for visNetwork graph)
```{r}
# create edge labels
# initialize edge labels list
edge_labels <- c()
# iterate with length of from_id_plus list
for (i in 1:length(from_id_plus)) {
  # create edge label and add to list
  edge_labels <- c(edge_labels, paste("Edge", from_id_plus[i], "-", to_id_plus[i]))
}
# create edge titles
# initialize edge titles list
edge_titles <- c()
# iterate with length of from_id_plus list
for (i in 1:length(from_id_plus)) {
  # create edge title and add to list
  edge_titles <- c(edge_titles, paste("Edge from", from_id_plus[i], "to", to_id_plus[i]))
}
# create edges data frame for Louvain algorithm applied graph
new_g_edges <- data.frame(
  from = from_id_plus,
  to = to_id_plus,
  label = edge_labels,
  arrows = "to",
  title = edge_titles
)
```

# graph using visNetwork
```{r}
#set.seed(11) (seed for similar looking graph)
set.seed(24)
visNetwork(new_g_nodes, new_g_edges, height = "850px", width = "100%") %>%
  visIgraphLayout() %>%
  visNodes(
    shape = "dot",
    color = list(
      background = "#0085AF",
      border = "#013848",
      highlight = "#FF8000"
    ),
    shadow = list(enabled = FALSE, size = 35)
  ) %>%
  visEdges(
    shadow = FALSE,
    color = list(color = "#0085AF", highlight = "#C62F4B")
  ) %>%
  visOptions(highlightNearest = list(enabled = T, degree = 1, hover = T),
             selectedBy = "group")
```

# graph using igraph
```{r}
set.seed(2)
plot(g)
set.seed(2)
plot(new_g, g)
```



### ANALYSIS

# CLOISTERED GRAPH

# get all connections between nodes
```{r}
# node connections list
node_connections <- c()
# iterate through all nodes in from_to list
for (i in seq(1, length(from_to), 2)){
  # append each to node to respective from node
  node_connections[[as.character(from_to[i])]] = c(node_connections[[as.character(from_to[i])]], from_to[i+1])
}
```

# find edges that belong to same group or different group and find all node connections in same group
```{r}
# nodes in same group list
group_node_connections <- list()
# edges that belong to a group
in_edge_groups <- list()
# edges that leave a group
out_edge_groups_f <- list()
# edges that enter a group
out_edge_groups_s <- list()

# iterate through all edges
for (i in 1:nrow(new_g_edges)){
  # get from node
  fnum <- new_g_edges$from[i]
  # get to node
  snum <- new_g_edges$to[i]
  # find which group fnum belongs to
  fnum_group <- new_g_nodes$group[which(new_g_nodes$id == fnum)]
  # find which group snum belongs to
  snum_group <- new_g_nodes$group[which(new_g_nodes$id == snum)]
  
  # if fnum and snum belong to same group
  if(fnum_group == snum_group){
    # add edge to in_edge_group
    in_edge_groups[[as.character(fnum_group)]] <- c(in_edge_groups[[as.character(fnum_group)]], sprintf('(%i, %i)', fnum, snum))
    # add nodes to group_node_connections
    group_node_connections[[as.character(fnum)]] <- c(group_node_connections[[as.character(fnum)]], snum)
  # fnum and snum don't belong to same group
  } else{
    # add edge to out_edge_group_f
    out_edge_groups_f[[as.character(fnum_group)]] <- c(out_edge_groups_f[[as.character(fnum_group)]], sprintf('(%i, %i)', fnum, snum))
    # add edges to out_edge_group_s
    out_edge_groups_s[[as.character(snum_group)]] <- c(out_edge_groups_s[[as.character(snum_group)]], sprintf('(%i, %i)', fnum, snum))
  }
}
```

# find MPNE for each group
```{r}
# max possible number of edges in a group
MPNE <- c()
# iterate through all groups
for (i in 1:length(new_g)) {
  # max possible number of edges calculated as n * (n-1) / 2 where n = number of nodes
  # append max possible number of edges to MPNE
  MNPE <- c(MPNE, (length(new_g[[i]])*(length(new_g[[i]])-1))/2)
}
```

# Cloistered Graph Analysis
```{r}
print("-------- Cloistered Graph Analysis --------")

# iterate over different groups
for (i in 1:length(MPNE)) {
  # Group i analysis
  print(sprintf("Group %i:", i))
  # find total number of edges entering and leaving
  edges_l_e <- length(out_edge_groups_f[[as.character(i)]]) + length(out_edge_groups_s[[as.character(i)]])
  # how many edges are in each group
  len_egroup <- length(in_edge_groups[[as.character(i)]])
  # how close group is to that of a complete graph as a percentage
  percent <-  len_egroup / (MPNE[[i]]) * 100
  # how many edges are missing from that of a complete graph
  missing_e <- MPNE[[i]] - len_egroup
  # group is cloistered graph or not
  cloistered <- FALSE
  
  # if number of edges leaving/entering group is 2 percent of max possible number of edges
  # then consider it a cloistered group
  if (edges_l_e < MPNE[[i]] * 0.02) {
    # print statement
    print(sprintf("   Group %i is a cloistered graph with %i edges leaving/entering the group", i, edges_l_e))
    # set cloistered to TRUE
    cloistered <- TRUE
  # if more than 2 percent then don't consider a cloistered graph
  } else {
    # print statement
    print(sprintf("   Group %i is not a cloistered graph with %i edges leaving/entering the group", i, edges_l_e))
  }
  
  # if percentage = 100% and graph is a cloistered graph then group is an ideal cloistered graph
  if (percent == 100 & cloistered & length(edges_l_e) == 0) {
    # print statement
    print(sprintf("   Group %i is an ideal cloistered graph", i))
  # if percentage = 100% but not cloistered graph then only complete graph
  } else if (percent == 100 & !cloistered){
    # print statement
    printf(sprintf("   Group %i is a complete graph", i))
  # if percentage != 100% and not a cloistered graph
  } else {
    # print statement
    print(sprintf("   Group %i is %f%% (%i/%i) of a complete graph with %i missing edges compared to that of a complete graph", i, percent, len_egroup, MPNE[[i]], missing_e))
  }
}
```



# TREE GRAPH

# function to create tree
```{r}
create_tree <- function(tree = list(), level = 1, node) {
  # if no nodes in current level of tree
  if (is.null(tree[[as.character(level)]])) {
    # add node to tree at level
    tree[[as.character(level)]] <- node
  # nodes exist in current level of tree
  } else {
    # add node to tree at level
    tree[[as.character(level)]] <- c(tree[[as.character(level)]], node)
  }
  
  # get children of  node
  children <- group_node_connections[[as.character(node)]]
  
  # if node has children
  if (!is.null(children)) {
    # iterate through children
    for (child in children) {
      # recursive call
      tree <- create_tree(tree, level + 1, child)
    }
  }

  # return tree
  return(tree)
}
```

# check for cycle function
```{r}
has_cycle <- function(node_list, group_node_connections) {
  # helper function for DFS
  dfs <- function(node) {
    
    # if a node has not been visited
    if (!visited[[node]]) {
      # mark node as visited
      visited[node] <- TRUE
      # add node to recursion stack
      rec_stack[node] <- TRUE
      
      # if the node is connected to other nodes in group
      if (!is.null(group_node_connections[[as.character(node)]])) {
        # iterate through other nodes in group
        for (near_node in group_node_connections[[as.character(node)]]) {
          # if the node has not been visited and recursion call on near_node
          if (!visited[[near_node]] && dfs(near_node)) {
            # return TRUE
            return (TRUE)
          # otherwise if the node near is on the recursion stack
          } else if (rec_stack[[near_node]]) {
            # return TRUE
            return (TRUE)
          }
          
        }
      }
    }
    # remove the node from recursion stack
    rec_stack[node] <- FALSE
    # return FALSE
    return (FALSE)
  }
  
  # list for visited nodes
  visited <- list()
  # list for recursion stack
  rec_stack <- list()
  # initialize visited and recursion stack
  for (node in node_list){
    visited[node] <- FALSE
    rec_stack[[node]] <- FALSE
  }
  
  # check each node in the group
  for (node in node_list) {
    # if dfs(node is true then there is a cycle)
    if (dfs(node)){
      # return TRUE
      return (TRUE)
    }
  }
  # there is no cycle so return FALSE
  return (FALSE)
}
```

# function to find statistics
```{r}
tree_statistics <- function(tree, group_node_connections){
  # set level error to 0
  level_error <- 0
  # set extra nodes to 0
  extra_nodes <- 0
  # set missing nodes to 0
  missing_nodes <- 0
  # set height balance to number of levels
  height_balance <- length(names(tree))
  # set nodes relative to ideal nodes to 0
  nrin <- 0
  # set total nodes to 1 (root node)
  total_nodes <- 1
  # set largest number of nodes in a level to 1 (root node)
  lnol <- 1
  # set number of nodes in a level to 1 (root node)
  nol <- 1
  
  # iterate through all levels in tree (skip level 0)
  for (level in names(tree)[-1]){
    # increase total nodes by number of nodes on level
    total_nodes <- total_nodes + length(tree[[as.character(level)]])
    # get level number of previous level
    prev_level <- as.character(as.numeric(level)-1)
    
    # if current number of nodes is greater than 2 ^ number of nodes from previous level
    if (length(tree[[as.character(level)]]) > 2^length(tree[[prev_level]])){
      # increase extra nodes by the difference
      extra_nodes <- extra_nodes + length(tree[[as.character(level)]]) - 2^length(tree[[prev_level]])
      # increase level error by 1
      level_error <- level_error + 1
    # if current number of nodes is less than 2 ^ number of nodes from previous level
    } else if (length(tree[[as.character(level)]]) < 2^length(tree[[prev_level]])){
      # increase missing nodes by difference
      missing_nodes <- missing_nodes + 2^length(tree[[prev_level]]) - length(tree[[as.character(level)]])
      # increase level error by 1
      level_error <- level_error + 1
    }
    
    # nol equal to number of nodes on current level
    nol <- length(tree[[as.character(level)]])
    # if nol less than or equal to lnol
    if (nol <= lnol){
      # decrease height_balance by 1
      height_balance <- height_balance - 1
    # if nol greater than lnol
    } else {
      # set lnol to nol
      lnol <- nol
    }
    
  }
  
  # set left_nodes to left half of children of root node
  left_nodes <- head(tree[['1']], floor(length(tree[['1']])/2))
  # set total left nodes to length of left_nodes
  total_left_nodes <- length(left_nodes)
  # set right_nodes to right half of children of root node
  right_nodes <- tail(tree[['1']], ceiling(length(tree[['1']])/2))
  # set total right nodes to length of right_nodes
  total_right_nodes <- length(right_nodes)
  
  # iterate through all nodes of left_nodes
  while (length(left_nodes) != 0){
    # if first node in left_nodes has children
    if (!is.null(group_node_connections[[as.character(left_nodes[1])]])){
      # increase total_left_nodes by number of children
      total_left_nodes <- total_left_nodes + length(group_node_connections[[as.character(left_nodes[1])]])
      # append children to left_nodes
      left_nodes <- c(left_nodes, group_node_connections[[as.character(left_nodes[1])]])
    }
    # pop off first node in left_nodes
    left_nodes <- left_nodes[-1]
  }
  # iterate through all nodes of right_nodes
  while (length(right_nodes) != 0){
    # if first node of right_nodes has children
    if (!is.null(group_node_connections[[as.character(right_nodes[1])]])){
      # increase total_right_nodes by number of children
      total_right_nodes <- total_right_nodes + length(group_node_connections[[as.character(right_nodes[1])]])
      # append children to right_nodes
      right_nodes <- c(right_nodes, group_node_connections[[as.character(right_nodes[1])]])
    }
    # pop off first node in right_nodes
    right_nodes <- right_nodes[-1]
  }
  
  # find nrin
  nrin <- sprintf("%i/%i", total_nodes, 2^(length(names(tree)))-1)
  
  # return statistics
  return (c(extra_nodes, missing_nodes, level_error, total_left_nodes, total_right_nodes, nrin, height_balance))
}
```

# function to find value of tree based on analysis
```{r}
tree_analysis <- function(temp_tree_statistic){
  left_nodes <- as.numeric(temp_tree_statistic[4])
  right_nodes <- as.numeric(temp_tree_statistic[5])
  # if left_nodes < right_nodes
  if (left_nodes < right_nodes) {
    # side_balance is ratio of left_nodes to right_nodes
    side_balance <- left_nodes / right_nodes
  # left_nodes > right_nodes
  } else{
    # side_balance is ratio of right_nodes to left_nodes
    side_balance <- right_nodes / left_nodes
  }
  
  # split nrin into numerator and denominator by /
  fraction <- strsplit(temp_tree_statistic[6], "/")[[1]]
  # get numerator of fraction
  numerator <- as.numeric(fraction[1])
  # get denominator of fraction
  denominator <- as.numeric(fraction[2])
  # calculate nrin
  nrin <- numerator / denominator
  
  node_error <- as.numeric(temp_tree_statistic[1]) + as.numeric(temp_tree_statistic[2])
  
  # node_error has weight of 0.35
  node_error_weight <- (1 - (node_error / denominator)) * 0.35
  # level_error has weight of 0.25
  level_error_weight <- (1 - (as.numeric(temp_tree_statistic[3]) / log((denominator+1), 2))) * 0.25
  # nrin has weight of 0.15
  nrin_weight <- nrin * 0.15
  # side_balance has weight of 0.15
  side_balance_weight <- side_balance * 0.15
  # height_balance has weight of 0.1
  height_balance_weight <- (1 - (as.numeric(temp_tree_statistic[7]) / log((denominator+1), 2))) * 0.1
  
  # sum up weights to calculate temp_tree_statistic and return max(0, weights)
  return (max(0, (node_error_weight + level_error_weight + side_balance_weight + height_balance_weight + nrin_weight)))
}
```

# Tree Graph Analysis
```{r}
print("-------- Tree Graph Analysis --------")

# create new from_to list for finding cycle
tree_cycle_graph_list <- list()
# iterate through nodes in same group
for (group in unique(new_g_nodes$group)){
  # add nodes in same group to tree_cycle_graph_list
  tree_cycle_graph_list[[as.character(group)]] <- new_g[[as.character(group)]]
}
     
# create potential root nodes list
potential_root_nodes <- list() 

# get each group in tree_cycle_graph_list
for (group in names(tree_cycle_graph_list)){
  # group must have more than 1 node
  if (length(tree_cycle_graph_list[[group]]) > 1) {
    # if there is a cycle
    if (has_cycle(tree_cycle_graph_list[[group]], group_node_connections)) {
      # print statement
      print(sprintf("   Group %s has a cycle so not a tree", group))
    # group does not have cycle
    } else {
      # look at edges in edges leaving groups
      for (edge in out_edge_groups_f[[group]]){
        # extract number from edge
        number <- gregexpr("[0-9]+", edge)
        result <- regmatches(edge, number)
        #get first number (the from node)
        node <- as.numeric((result[[1]][1]))
        # add node to potential root nodes list
        potential_root_nodes[[group]] <- unique(c(potential_root_nodes[[group]], node))
      }
      # look at edges in edges entering groups
      for (edge in out_edge_groups_s[[group]]){
        # extract number from edge
        number <- gregexpr("[0-9]+", edge)
        result <- regmatches(edge, number)
        #get first number (the from node)
        node <- as.numeric((result[[1]][2]))
        # add node to potential root nodes list
        potential_root_nodes[[group]] <- unique(c(potential_root_nodes[[group]], node))
      }
      
      # if there are no nodes for the group in potential_root_nodes list
      if (length(potential_root_nodes[[group]]) == 0){
        # add every node in the group to potential_root_nodes list
        for (node in tree_cycle_graph_list[[group]]){
          # only keep unique nodes
          potential_root_nodes[[group]] <- unique(c(potential_root_nodes[[group]], node))
        }
      }
      
    }
  }
}
  
# create best root node analysis list (keeps track of best root node, best tree value, and best tree statistics)
broot_node_analysis <- list()

# group index
group_i <- 1
# iterate through groups in potential_root_nodes list
for (group in names(potential_root_nodes)){
  # statistic of best tree in group
  best_tree_statistic <- c()
  # analysis of best tree in group
  best_tree_analysis <- 0
  # root node in group that led to best tree
  best_node <- 0
  # iterate through nodes in groups in potential_root_nodes
  for (node in potential_root_nodes[[group]]){
    # create a tree for root node
    tree <- create_tree(list(), 1, node)
    
    # get statistics of temp_tree
    temp_tree_statistic <-  tree_statistics(tree, group_node_connections)
    # get analysis of temp_tree
    temp_tree_analysis <- tree_analysis(temp_tree_statistic)
    
    # if temp_tree analysis >= best_tree_analysis
    if (temp_tree_analysis >= best_tree_analysis){
      # set best_tree_statistic to temp_tree_statistic
      best_tree_statistic <- temp_tree_statistic
      # set best_tree_analysis to temp_tree_analysis
      best_tree_analysis <- temp_tree_analysis
      # set best_node to node
      best_node <- node
    }
  }
  # save best_tree_statistic in broot_node_analysis
  broot_node_analysis[[group_i]] <- c(best_node, best_tree_analysis, best_tree_statistic)
  
  # Group i analysis
  print(sprintf("Group %s:", group))
  # if the tree statistic for the group is greater than or equal to 0.8 then the group is a tree
  if (best_tree_analysis >= 0.7){
    # print statements
    print(sprintf("   Group %s is a tree graph with node %i being the root node", group, as.numeric(broot_node_analysis[[group_i]][1])))
    print(sprintf("   %f%% of the group shares the same properties as those of an ideal tree graph", 100*as.numeric(broot_node_analysis[[group_i]][2])))
    print(sprintf("   %i Node Errors (%i Extra Nodes, %i Missing Nodes), %i Level Errors, %i Left Nodes and %i Right Nodes, %s (%f%%) NRIN, %i Height Balance", as.numeric(broot_node_analysis[[group_i]][3])+as.numeric(broot_node_analysis[[group_i]][4]), as.numeric(broot_node_analysis[[group_i]][3]), as.numeric(broot_node_analysis[[group_i]][4]), as.numeric(broot_node_analysis[[group_i]][5]), as.numeric(broot_node_analysis[[group_i]][6]), as.numeric(broot_node_analysis[[group_i]][7]), broot_node_analysis[[group_i]][8], 100*as.numeric(eval(parse(text = broot_node_analysis[[group_i]][8]))),  as.numeric(broot_node_analysis[[group_i]][9])))
  # if the tree statistic for the group is less than 0.8 then the group is not a tree
  }else{
    # print statements
    print(sprintf("   Group %s is not a tree and has a structure closest to a tree graph when node %i is the root node", group, as.numeric(broot_node_analysis[[group_i]][1])))
    print(sprintf("   %f%% of the group sharing the same properties as those of an ideal tree graph", 100*as.numeric(broot_node_analysis[[group_i]][2])))
    print(sprintf("   %i Node Errors (%i Extra Nodes, %i Missing Nodes), %i Level Errors, %i Left Nodes and %i Right Nodes, %s (%f%%) NRIN, %i Height Balance", as.numeric(broot_node_analysis[[group_i]][3])+as.numeric(broot_node_analysis[[group_i]][4]), as.numeric(broot_node_analysis[[group_i]][3]), as.numeric(broot_node_analysis[[group_i]][4]), as.numeric(broot_node_analysis[[group_i]][5]), as.numeric(broot_node_analysis[[group_i]][6]), as.numeric(broot_node_analysis[[group_i]][7]), broot_node_analysis[[group_i]][8], 100*as.numeric(eval(parse(text = broot_node_analysis[[group_i]][8]))),  as.numeric(broot_node_analysis[[group_i]][9])))
  }
  # increase group_i by 1
  group_i <- group_i + 1
}
```



#LINE BRANCH GRAPH

# function to find longest line in group
```{r}
longest_line <- function(node, group_node_connections, visited = c()) {
  # if the initial node has been visited
  if (node %in% visited){
    # return length of 0 and an empty path
    return(list(length = 0, path = c()))
  } 
  
  # initial node has not been visited so add node to visited
  visited <- c(visited, node)
  
  # if node does not have connections to other nodes
  if (is.null(group_node_connections[[as.character(node)]])) 
    # return length of 1 (just initial node) and the node as the path
    return(list(length = 1, path = c(node)))
  
  # longest line length
  max_length <- 0
  # path of longest line
  max_path <- c()
  
  # explore all nodes connected to node
  for (near_node in group_node_connections[[as.character(node)]]) {
    # recursive call
    line <- longest_line(near_node, group_node_connections, visited)
    # if the length of the line is greater than max_length
    if (line$length > max_length) {
      # set max_length to the line's length
      max_length <- line$length
      # set max_path to the line's path
      max_path <- line$path
    }
  }
  
  # return max_length + 1 to include current node and max_path
  return(list(length = max_length + 1, path = c(node, max_path)))
}
```

# create reverse group node connections list
```{r}
# function to create reverse group node connections
reverse_connections <- function(connections) {
  # initialize reverse group node connections list
  reversed <- list()
  # iterate through all from nodes in connections list (from nodes: nodes that point to other nodes)
  for (from in names(connections)) {
    # iterate through all to nodes (to nodes: nodes that are pointed to)
    for (to in connections[[from]]) {
      # if to node is not in the names of reversed list
      if (!as.character(to) %in% names(reversed)) {
        # initialize nodes pointed at to node in reversed list
        reversed[[as.character(to)]] <- c()
      }
      # add from node to reversed list at index to node
      reversed[[as.character(to)]] <- c(reversed[[as.character(to)]], as.numeric(from))
    }
  }
  # return reversed list
  return(reversed)
}

# create reversed connections
reversed_gn_connections <- reverse_connections(group_node_connections)
```

# function to find statistics
```{r}
lb_statistics <- function(line, group_nodes, group_node_connections){
  # set branch_nodes to number of nodes connected to last node in line
  branch_nodes <- group_node_connections[[as.character((tail(line, n=1)))]]

  # set line_error to (number of nodes in group) - ((number of nodes in line) + (branching nodes))
  line_error <- length(group_nodes) - (length(line) +  length(branch_nodes))
  # set llrill to (number of nodes in line) / (2*log(length(branch_nodes) + 1) + 5)
  llrill <- sprintf("%i/%i", length(line), as.integer((2*log(length(branch_nodes) + 1) + 5)))
  
  # initialize branch_continuation to 0
  branch_continuation <- 0
  # iterate through all branch_nodes
  for (node in branch_nodes){
    # add number of nodes connected to node to branch_continuation
    branch_continuation <- branch_continuation + length(group_node_connections[[as.character(node)]])
  }
  
  # initialize branch_positions
  branch_positions <- c()
  # set line_index to 0
  line_index <- 0
  
  # iterate through nodes in line
  for (node in line[-length(line)]){
    # increase line_index by 1
    line_index <- line_index + 1
    # if the node has more than 1 connection in group
    if (length(group_node_connections[[as.character(node)]]) > 1){
      # add position relative to line to branch_positions
      branch_positions <- c(branch_positions, sprintf("%i/%i", line_index, length(line)))
      
    }
  }
  if (length(branch_positions) == 0){
    branch_positions = 0
  }
  
  # return statistics
  return (c(line_error, llrill, branch_continuation, branch_positions))
}
```

# function to find value of line branch based on analysis
```{r}
lb_analysis <- function(lb_statistic){
  # split llrill into numerator and denominator by /
  fraction <- strsplit(lb_statistic[2], "/")[[1]]
  # get numerator of fraction
  line_length <- as.numeric(fraction[1])
  # get denominator of fraction
  i_line_length <- as.numeric(fraction[2])
  # calculate llrill
  llrill <- line_length / i_line_length
  
  # calculate how many branch nodes there are in group (isolate branch_nodes in ideal line length = 2*log(length(branch_nodes) + 1) + 5)
  branch_nodes <- exp((i_line_length - 5)/2)-1
  # calculate line_error such that line_error is relative to number of nodes in ideal line length + branching nodes
  line_error <- as.integer(lb_statistic[1]) / (i_line_length + as.integer(branch_nodes))
  
  # calculate branch_continuation relative to number of nodes in line
  branch_continuation <- as.integer(lb_statistic[3]) / line_length
  
  # initialize penalty_sum
  penalty_sum <- 0
  # iterate through positions of branching nodes in line relative to line
  for (position_r in lb_statistic[4]){
    # split position relative into numerator and denominator by /
    fraction <- strsplit(position_r, "/")[[1]]
    # get position (numerator)
    position <- as.numeric(fraction[1])
    # add ((number of nodes in line) - (position of branching node in line)) / (number of nodes in line) to penalty_sum
    penalty_sum <- penalty_sum + ((line_length - position) / line_length)
  }
  
  # calculate branch_position (1 - penalty_sum / (number of branching nodes in line))
  branch_positions <- 1 - penalty_sum/length(lb_statistic[4])
  
  # line_error has weight of 0.4
  line_error_weight <- (1 - line_error) * 0.4
  # llrill has weight of 0.3
  llrill_weight <- llrill * 0.3
  # branch_continuation has weight of 0.2
  branch_continuation_weight <- (1 - branch_continuation) * 0.2
  # branch_positions has weight of 0.1
  branch_positions_weight <- branch_positions * 0.1
  
  # sum up weights to calculate lb_statistic and return max(0, weights)
  return (max(0, (line_error_weight + llrill_weight + branch_continuation_weight + branch_positions_weight)))
}
```

# LINE BRANCH GRAPH ANALYSIS
```{r}
print("-------- Line Branch Graph Analysis --------")

# create new from_to list for finding cycle
lb_cycle_graph_list <- list()
# iterate through nodes in same group
for (group in unique(new_g_nodes$group)){
  # add nodes in same group to lb_cycle_graph_list
  lb_cycle_graph_list[[as.character(group)]] <- new_g[[as.character(group)]]
}
    
# create potential end nodes list
potential_end_nodes <- list()  

# get each group in lb_cycle_graph_list
for (group in names(lb_cycle_graph_list)){
  # group must have more than 1 node
  if (length(lb_cycle_graph_list[[group]]) > 1) {
    # if there is a cycle
    if (has_cycle(lb_cycle_graph_list[[group]], group_node_connections)) {
      # print statement
      print(sprintf("   Group %s has a cycle so not a line branch", group))
    # group does not have cycle
    } else {
      potential_end_nodes[[group]] <- lb_cycle_graph_list[[group]]
    }
  }
}

# create best start node analysis list (keeps track of best start node, best line branch value, and best line branch statistics)
bstart_path <- list()

# iterate through groups in potential_end_nodes list
for (group in names(potential_end_nodes)){
  # path of longest line
  temp_longest_path <- c()
  # most branching nodes
  largest_branch <- 0
  # longest path with most branching
  longest_path <- list()
  # iterate through nodes in groups in potential_end_nodes
  for (node in potential_end_nodes[[group]]){
    # if the node has more branches than largest_branch
    if (length(group_node_connections[[as.character(node)]]) > largest_branch){
      # set largest_branch to the number of branches of the node
      largest_branch <- length(group_node_connections[[as.character(node)]])
      # set the longest_path to the longest line of nodes in the group that ends with node
      longest_path <- longest_line(node, reversed_gn_connections)$path
    # if the node has equal number of branches to largest_branch
    } else if (length(group_node_connections[[as.character(node)]]) == largest_branch){
      # find the longest line of nodes in the group that ends with node
      temp_longest_path <- longest_line(node, reversed_gn_connections)$path
      # if the length of this line is greater than the length of longest_path
      if (length(temp_longest_path) > length(longest_path)){
        # set longest_path to the line
        longest_path <- temp_longest_path
      }
    }
  }
  # save path using best start node to bstart_path
  bstart_path[[group]] <- c(rev(longest_path))
}

bstart_node_analysis <- list()
# iterate through groups in potential_end_nodes list
for (group in names(potential_end_nodes)){
  # get line branch statistic using path in bstart_path at the group as the start node
  lb_statistic <- lb_statistics(bstart_path[[group]], lb_cycle_graph_list[[group]], group_node_connections)
  # get line branch value from lb_analysis
  lb_value <- lb_analysis(lb_statistic)
  # add line branch value and lb_statistic to bstart_node_analysis
  bstart_node_analysis[[group]] <- c(lb_value, lb_statistic, bstart_path[[group]])
  # Group i analysis
  print(sprintf("Group %s:", group))
  # if the line branch analysis for the group is greater than or equal to 0.8 then the group is a line branch
  if (lb_value >= 0.7){
    # print statements
    print(sprintf("   Group %s is a line branch with node %i being the start node", group, as.numeric(bstart_node_analysis[[group]][6])))
    last_i <- length(bstart_node_analysis[[group]])
    print(sprintf("   The line is from node %s", paste(as.character(as.integer(bstart_node_analysis[[group]][6:last_i])), collapse=" to ")))
    print(sprintf("   Node %s then branches into nodes %s", as.character(as.integer(bstart_node_analysis[[group]][last_i])), paste(as.character(group_node_connections[[as.character(bstart_node_analysis[[group]][last_i])]][1:length(group_node_connections[[as.character(bstart_node_analysis[[group]][last_i])]])]), collapse=", ")))
    print(sprintf("   %f%% of the group shares the same properties as those of an ideal line branch", 100*as.numeric(bstart_node_analysis[[group]][1])))
    print(sprintf("   %i Line Errors, %s (%f%%) LLRBN, %i Branch Continuation, %s Branch Positions", as.numeric(bstart_node_analysis[[group]][2]), bstart_node_analysis[[group]][3], 100*as.numeric(eval(parse(text = bstart_node_analysis[[group]][3]))), as.numeric(bstart_node_analysis[[group]][4]), bstart_node_analysis[[group]][5]))
  # if the line branch analysis for the group is less than 0.8 then the group is not a line branch
  }else{
    # print statements
    print(sprintf("   Group %s is not a line branch and has a structure closest to a line branch when node %i is the start node", group, as.numeric(bstart_node_analysis[[group]][6])))
    last_i <- length(bstart_node_analysis[[group]])
    print(sprintf("   The line is from node %s", paste(as.character(as.integer(bstart_node_analysis[[group]][6:last_i])), collapse=" to ")))
    print(sprintf("   Node %s then branches into nodes %s", as.character(as.integer(bstart_node_analysis[[group]][last_i])), paste(as.character(group_node_connections[[as.character(bstart_node_analysis[[group]][last_i])]][1:length(group_node_connections[[as.character(bstart_node_analysis[[group]][last_i])]])]), collapse=", ")))
    print(sprintf("   %f%% of the group shares the same properties as those of an ideal line branch", 100*as.numeric(bstart_node_analysis[[group]][1])))
    print(sprintf("   %i Line Errors, %s (%f%%) LLRBN, %i Branch Continuation, %s Branch Positions", as.numeric(bstart_node_analysis[[group]][2]), bstart_node_analysis[[group]][3], 100*as.numeric(eval(parse(text = bstart_node_analysis[[group]][3]))), as.numeric(bstart_node_analysis[[group]][4]), bstart_node_analysis[[group]][5]))
  }
}
```
